---
title: "Dog survey analysis"
output: html_notebook
---

```{r}
library(tidyverse)
library(here)

dogs_clean <- read_csv(here("clean_data/dogs_clean.csv"))
```

# 1. The client only counts a valid email address as one ending in ‘.com’. How many survey results have a valid email address?
```{r}
dogs_clean %>% 
  filter(str_detect(email, ".com$")) %>% 
  count()
```

181 valid email addresses

# 2. What’s the average amount spent on dog food for each dog size?
```{r}
dogs_clean %>% 
  group_by(dog_size) %>% 
  summarise(mean_spent = mean(amount_spent_on_dog_food, na.rm = T))
```



# 3. For owners whose surname starts with a letter in the second half of the alphabet (N onwards) what is the average age of their dog?
```{r}
dogs_clean %>% 
  filter(str_detect(last_name, "[NOPQRSTUVWXYZ]")) %>% 
  summarise(mean_age = mean(dog_age, na.rm = T))
```



# 4. The dog_age column is the age in dog years. If the conversion is 1 human year = 6 dog years, then what is the average human age for dogs of each gender?

```{r}
dogs_clean %>% 
  group_by(dog_gender) %>% 
  summarise(mean_age_human_years = mean(human_age, na.rm = T))
```


# 5. Create a plot of results of question 4.

```{r}
dogs_clean %>% 
  group_by(dog_gender) %>% 
  #summarise(mean_age_human_years = mean(human_age, na.rm = T)) %>% 
  ggplot(aes(x = dog_gender, y = human_age)) +
  geom_boxplot() +
  geom_jitter()
```

