---
title: "Candy data task"
output: html_notebook
---

```{r}
library(tidyverse)
library(here)

candy_data <- read_csv(here("clean_data/candy_clean.csv"))

# candy_data <- candy_data %>% 
#   mutate(rating = as.factor(rating),
#          going_out = as.factor(going_out),
#          gender = as.factor(gender))
```


Q1. What is the total number of candy ratings given across the three years. (Number of candy ratings, not the number of raters. Don’t count missing values)
```{r}
candy_data %>% 
  select(rating) %>% 
  filter(!is.na(rating)) %>% 
  summarise(n_ratings = n())
```
Answer: 757108 ratings


Q2. What was the average age of people who are going out trick or treating?
```{r}
candy_data %>% 
  select(age_years, going_out) %>% 
  filter(going_out == "Yes") %>% 
  summarise(av_age = mean(age_years, na.rm = TRUE))
```
Average age of people going out is 35.1

Q3. What was the average age of people who are not going trick or treating?
```{r}
candy_data %>% 
  select(age_years, going_out) %>% 
  filter(going_out == "No") %>% 
  summarise(av_age = mean(age_years, na.rm = TRUE))
```
Average age of people not going out is 39.23


Q4. For each of joy, despair and meh, which candy bar received the most of these ratings?
```{r}
## good way
candy_data %>% 
  filter(!is.na(rating)) %>% 
  group_by(rating, candy) %>% 
  summarise(
    count = n()
  ) %>% 
  filter(count == max(count))

## slow way
# joy
candy_data %>% 
  select(candy, rating) %>% 
  filter(rating == "joy") %>% 
  group_by(candy) %>% 
  summarise(N_joy = n()) %>% 
  arrange(desc(N_joy)) %>% 
  head(1)
  
# despair
candy_data %>% 
  select(candy, rating) %>% 
  filter(rating == "despair") %>% 
  group_by(candy) %>% 
  summarise(N_despair = n()) %>% 
  arrange(desc(N_despair)) %>% 
  head(1)

# meh
candy_data %>% 
  select(candy, rating) %>% 
  filter(rating == "meh") %>% 
  group_by(candy) %>% 
  summarise(N_meh = n()) %>% 
  arrange(desc(N_meh)) %>% 
  head(1)
```

Any full sized bar, glowsticks, and lollipops, respectively


Q5. How many people rated Starburst as despair?
```{r}
candy_data %>% 
  select(candy, rating) %>% 
  filter(rating == "despair",
         candy == "starburst") %>% 
  summarise(N = n()) 
```
1990 people rated it despair

Q6. For the next three questions, count despair as -1, joy as +1, and meh as 0.

```{r}
candy_data  <-  candy_data %>% 
  mutate(rating_num = case_when(
    rating == "despair" ~ -1,
    rating == "joy" ~ 1,
    rating == "meh" ~ 0
  )
  )

```



Q7. What was the most popular candy bar by this rating system for each gender in the dataset ?
```{r}
## good way
candy_data %>% 
  group_by(gender, candy) %>% 
  summarise(
    avg_rating = mean(rating_num, na.rm = TRUE)
  ) %>% 
  filter(avg_rating == max(avg_rating))


## slow way
candy_data %>% 
  select(candy, rating_num, gender) %>% 
  group_by(gender, candy) %>% 
  summarise(count =sum(rating_num, na.rm = TRUE)) %>% 
  arrange(desc(count)) %>% 
  filter(gender == "Male",
         count == max(count))

candy_data %>% 
  select(candy, rating_num, gender) %>% 
  group_by(gender, candy) %>% 
  summarise(count =sum(rating_num, na.rm = TRUE)) %>% 
  arrange(desc(count)) %>% 
  filter(
         count == max(count))

candy_data %>% 
  select(candy, rating_num, gender) %>% 
  group_by(gender, candy) %>% 
  summarise(count =sum(rating_num, na.rm = TRUE)) %>% 
  arrange(desc(count)) %>% 
  filter(count == max(count))
```



Q8. What was the most popular candy bar in each year?
```{r}
# mean ratings 
candy_data %>% 
  group_by(year, candy) %>% 
  summarise(
    avg_rating = mean(rating_num, na.rm = TRUE)
  ) %>% 
  filter(avg_rating == max(avg_rating))

# summed ratings (because 1s + 0s + -1s will give overall ranking)
candy_data %>% 
  group_by(year, candy) %>% 
  summarise(
    sum_rating = sum(rating_num, na.rm = TRUE)
  ) %>% 
  filter(sum_rating == max(sum_rating))
```

Any full sized candy bar year on year, duh.

Q9. What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?
```{r}
candy_data %>% 
  mutate(country = if_else(country %in% c("united states", "united kingdom", "canada"), country, "other")) %>%  # mutate country column to either show the big 3 countries, else just 'other'
  group_by(country, candy) %>% 
  summarise(
    avg_rating = mean(rating_num, na.rm = TRUE) # give average rating per country and candy
  ) %>% 
  filter(avg_rating == max(avg_rating)) %>% # filter down to just the top average rated candy per country
  arrange(desc(country)) # descending order just for aesthetics
```

